---
title: "Gapminder_analyses"
author: "Benjamin Weiner"
date: "10/4/2015"
output: html_document
---

## Demographic analyses with R

These are analyses of the Gapminder demographic data using Rstudio.

The document and plots are produced using R markdown language.


```{r load-dplyr, echo=FALSE}
#install.packages("dplyr")
library("dplyr")
```

```{r read-in-data}
gap.in <- read.table("output/combined_gapMinder.tsv", sep="\t", header=TRUE)

gap.in %>%
  select(pop,country) %>%
  head()
gap.in %>%
  select(year,pop,country, continent) %>%
  filter(year == 2002) %>%
  group_by(continent) %>%
  summarise(mean = mean(pop))


```

``` {r define-plot}

make_plot <- function(countryname) {
  foo <- gap.in %>%
    select(year, country, pop, gdpPercap) %>%
    filter(country == countryname) 
  #print(foo)
  plot(x = foo$year, y = foo$gdpPercap,
         xlab = "Year",
         ylab = "GDP per capita",
         main = paste("GDP for", countryname),
         #cex = foo$pop/1e7,
         col = "red")
}

```


``` {r plot-japan}

make_plot("Japan")

```

Japan's GDP shows substantial growth after WWII until a period of stagnation beginning in the 1990s.


``` {r plot-southkorea}

make_plot("Korea, Rep.")

```

South Korea's GDP was relatively flat until the late 1970s and has increased dramatically since.

``` {r plot-chile}

make_plot("Chile")

```

Chile's GDP plot shows slow growth and an actual decline in the late 1970s, probably related to the coup and military government, until beginning a steep incrase in the 1990s after the return to democratic government.





Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot. You can interpret code inline like r (2+2) = `r (2+2)`.
