---
title: "Gapminder_analyses"
author: "Benjamin Weiner"
date: "10/4/2015"
output: html_document
---

## Demographic analyses with R

These are analyses of the Gapminder demographic data using Rstudio.

The document and plots are produced using R markdown language.


```{r load-dplyr, echo=FALSE}
#install.packages("dplyr")
library("dplyr")
```

```{r read-in-data, echo=FALSE}
gap.in <- read.table("output/combined_gapMinder.tsv", sep="\t", header=TRUE)

# gap.in %>%
#  select(pop,country) %>%
#  head()
# gap.in %>%
#  select(year,pop,country, continent) %>%
#  filter(year == 2002) %>%
#  group_by(continent) %>%
#  summarise(mean = mean(pop))


```

``` {r define-plot, echo=FALSE}

make_plot <- function(countryname) {
  foo <- gap.in %>%
    select(year, country, pop, gdpPercap) %>%
    filter(country == countryname) 
  #print(foo)
  plot(x = foo$year, y = foo$gdpPercap,
         xlab = "Year",
         ylab = "GDP per capita",
         main = paste("GDP for", countryname),
         #cex = foo$pop/1e7,
         col = "red")
}

```


``` {r plot-japan, echo=FALSE}

make_plot("Japan")

```

Japan's GDP shows substantial growth after WWII until a period of stagnation beginning in the 1990s.


``` {r plot-southkorea, echo=FALSE}

make_plot("Korea, Rep.")

```

South Korea's GDP was relatively flat until the late 1970s and has increased dramatically since.

``` {r plot-chile, echo=FALSE}

make_plot("Chile")

```

Chile's GDP plot shows slow growth and an actual decline in the late 1970s, probably related to the coup and military government, until beginning a steep incrase in the 1990s after the return to democratic government.

``` {r lifeterms, echo=FALSE}

# This is mean lifeExp across countries, not weighted properly by people
#gap.in %>%
#  select(year,pop,country, continent, lifeExp) %>%
#    group_by(continent) %>%
#  summarise(mean = mean(lifeExp))

# This should be mean lifeExp weighted per person
avglife <- gap.in %>%
  select(year,pop,country, continent, lifeExp) %>%
    group_by(continent) %>%
  summarise(mean = sum(lifeExp*pop) / sum(pop))

avggdp <- gap.in %>%
  select(year,pop,country, continent, lifeExp, gdpPercap) %>%
    group_by(continent) %>%
  summarise(mean = sum(gdpPercap*pop) / sum(pop))

print(avglife)

print(avggdp)

plot(avggdp,avglife,
     xlab = "Continent average GDP",
     ylab = "Continent average life expectancy")

```

``` {r print_avg}

print(avglife)

print(avggdp)
```

The mean life expectancy varies greatly from continent to continent and is correlated with GDP per capita.

``` {r lifegdp, echo=FALSE}

plot(log10(gap.in$gdpPercap), gap.in$lifeExp,
     xlab = "Country log10 GDP per capita",
     ylab = "Country average life expectancy")

```

On a country by country basis there is a strong correlation between log GDP per capita and life expectancy, with some possibly interesting low outliers.

``` {r lifehist, echo=FALSE}

hist(gap.in$lifeExp,
     xlab='Years',
     ylab='Number',
     main='Life expectancy per country')

hist(gap.in$lifeExp,
     breaks=25,
     xlab='Years',
     ylab='Number',
     main='Life expectancy per country')

```

The life expectancy histogram is skewed, with a long tail towards short life expectancy.  The finer-scale histogram shows that the tail is not steadily declining, but there is a plateau, or even a second peak around 40-50 years, and then a strong decline with few countries below 40 years.

